Active Learning for SAT-Solver Benchmarking

### 2022-09-26 paper/paper.tex
Add comments from PdF paper, part 1 (Sections 1-3)
... and slightly adapt text in some places (comments I directly resolved or
other stuff I noticed).

\begin{abstract}
  Experimentation is one of the main phases in algorithm engineering, including the archetypal $\mathcal{NP}$-complete propositional-sat\-is\-fia\-bi\-li\-ty problem (SAT).
  While the field of SAT-solver benchmarking is well established, traditional benchmark-set-selection approaches do not optimize for benchmark runtime.
  Benchmark-set selection is the process of choosing a subset of instances out of a pool to be representative and reliable in discriminating SAT solvers.
  This paper presents an active-learning-based benchmark-set-selection approach that solves the new-solver problem:
  Given a new solver, we want to find its ranking amidst its competitors with as little runtime as possible while still giving accurate ranking predictions.
  To do so, we use well-known techniques from the domain of active learning.
  We can determine a new solver's rank regarding the SAT Competition 2022 Anniversary Track instances in about \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset.
  We achieve this while still maintaining a prediction accuracy of about \SI{92}{\%} regarding the PAR-2 ranking of this solver if it were evaluated on the complete dataset.
  Instance-selection decisions made by our approach also allow for discussing instance-family importance in the selection process. 
  Our tool provides the engineer of propositional-satisfiability solvers with a faster way of determining a new SAT-solver's performance in comparison to sampling the benchmark set by hand.

@@

\section{Introduction}
\label{sec:intro}
One of the main phases of algorithm engineering is experimentation.
This is al\-so true for the archetypal $\mathcal{NP}$-complete pro\-po\-si\-tio\-nal-sat\-is\-fia\-bi\-li\-ty problem (SAT).
While state-of-the-art solvers embark upon more and more problem instances, SAT-solver evaluation is steadily getting more costly.
Usually, submissions to competitions like the annual SAT~Competition~\cite{FroleyksHIJS21} themselves precede heavy experimentation and are the remnant of a multitude of ideas and considerations.
While the field of SAT-solver benchmarking is well established, traditional benchmark-set-selection approaches~\cite{Gelder11,HoosKSS13} do not optimize for benchmark runtime.
Their goal is to select a representative and robust benchmark set for discriminating solvers reliably.
Within our work, we show that we can arrive at the same conclusions with high accuracy in only a fraction of the time that would be needed to evaluate the full benchmark set.
To be more specific, we address the following problem.

@@

We want to find a scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ assigning each solver a score.
Having a set of solvers with their respective scores induces a ranking.
This allows for both comparing the solvers as well as identifying solvers with similar behavior.
While the annual SAT~Competitions~\cite{FroleyksHIJS21} rely on the PAR-2 score ranking, i.e, the average runtime with a penalty timeout-term of $2\tau$, which we want to predict, this work examines a model-based way of ranking solvers.

This current article is proposing a novel active-learning-based benchmark-set-selection approach.
Our algorithm is unique in that it is the only one, to the best of our knowledge, to address all of the following features (also refer to Tab.~\ref{tab:requirements}).
Our benchmarking process outputs a \textit{ranking} to compare the new solver against the existing set of solvers as described above.
We optimize the \textit{runtime} that our strategy needs to arrive at its conclusion.
We use problem-instance and runtime \textit{features} $\varphi\!: \mathcal{A} \times \mathcal{I} \rightarrow \mathbb{R}^n$ since it already has proven to be helpful in classic benchmark-set-selection approaches~\cite{HoosKSS13} as well as in portfolio selection~\cite{CollauttiMMO13,NgokoCT19}.
Moreover, we select instances \textit{non-randomly} and \textit{incrementally}.
By doing so, we have a greater impact on the properties of the benchmarking approach such as its required runtime.
Sampling instances incrementally means taking runtime information from already done experiments into account when choosing the next.
Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \textit{scoring function} that shows by which margin a solver is worse and how similar it is to existing approaches (\textit{interpretability}).
Finally, our strategy's cost \textit{scales linearly} with the cardinality of the set of existing solvers $\mathcal{A}$ to compare against and the instances $\mathcal{I}$ within our pool of possible benchmark instances.
We avoid comparing pairs of solvers.
This would scale quadratically.

To sum up our contribution, we take an existing benchmark set, e.g., the SAT Competition~2022 Anniversary Track~\cite{sat2022}, and predict a new solver's PAR2 score on this dataset with only a fraction of the runtime that would be needed to evaluate this solver on the complete set of instances.

For reproducibility, our source code, all experiments, and data are available on GitHub\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}}.

@@

Tab.~\ref{tab:requirements} shows a concise overview of existing approaches with a related goal.
While traditional benchmark-set-selection~\cite{HoosKSS13} lacks the aspect of runtime optimization, related work within algorithm configuration~\cite{HutterHL11} is concerned with finding better configurations fast.
However, algorithm configuration is not suited for ranking.
There is already existing work employing active learning for SAT-solver discrimination~\cite{MatriconAFSH21}.
This approach has some major shortcomings though.
Finally, our approach seeks to provide all stated features.
Note that all approaches already make use of problem-instance features as it showed to be beneficial in past work~\cite{HoosKSS13,CollauttiMMO13,NgokoCT19}.

\subsubsection{Traditional benchmark-set-selection.}
Recent studies~\cite{abs-2107-07002,NiesslHWCB22} show that selecting a benchmark set in general is a difficult endeavor for multiple reasons.
Dehghani~et.~al.~\cite{abs-2107-07002} describe how a biased benchmarking set can easily lead to fallacious interpretations.
Moreover, Nießl~et.~al.~\cite{NiesslHWCB22} state that benchmark-set design has many movable parts, e.g., performance measures, aggregation, and handling of missing values.
Questionable research practices might modify those design elements a-posteriori to fit expectations, leading to biased results.

Hoos et. al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting a set of SAT benchmark instances, for instance for a competition, e.g., the SAT Competition~2020~\cite{FroleyksHIJS21}.
The selection strategy presented is static, i.e., it does not depend on the solvers to distinguish.
Desirable properties are a variety of instances (avoids over-fitting of solver approaches), adapted instance hardness (not too easy but also not too hard; solvers cannot be distinguished if too many time-out or finish immediately), and no duplicates (having duplicates gives those instances more weight).
To avoid having too similar instances, they make use of a distance-based approach using the SATzilla features~\cite{XuHHL08,features}.
Those features comprise general instance-size features and graph-representation features among others.
Their approach \textit{ranks} solvers by providing its PAR-2 score, is \textit{feature-based}, and is \textit{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
Also, their approach is \textit{interpretable} in the sense that the obtained PAR-2 score induces a distance between solvers.
However, they do not optimize for benchmark \textit{runtime} and select instances \textit{randomly}\footnote{Random with constraints on the instance hardness and feature distance.}.

@@

SMAC~\cite{HutterHL11} tries to find a close-to-optimal configuration for a given SAT-solver on a given set of instances.
Although this task is different from our goal, e.g., we do not need to navigate configuration space, they use a similar framework.
There are mainly two similarities.
First, both approaches are model-based.
While SMAC tries to gradually build a model of the solver's performance as a function of the configuration parameters, we want to gradually build a model of a new solver's performance based on already known runtimes of instances and runtime information on already known solvers.
Second, both approaches employ some kind of selection strategy that is based upon the previously mentioned model.
Theirs selects configurations rather than instances though.
Furthermore, Hutter et. al.~\cite{HutterHL11} use an approach that makes use of exploitation (what configuration parameters are already good) and exploration (parameter space with little data points) to select the next configuration for which to acquire runtime information. SMAC's configuration sampling has also proven to be better than random sampling in all tested scenarios and has even proven to be significantly better in most cases~\cite{HutterHL11}.

While this approach considers the configuration's \textit{runtime}, is \textit{feature-based}, and \textit{scales linearly} by having a fixed time budget, it is not capable of the other three features listed in Tab.~\ref{tab:requirements}.
First, an algorithm configurator cannot be used to \textit{rank} a new solver since it only looks at promising solvers/configurations. If the new solver does not belong to this category (new ideas are seldomly the best approach right from the start, i.e., they need some tweaking), we do not have any information on how this solver performed.
Second, while using a model-based selection strategy to sample configurations, instance sampling is done \textit{randomly}.
And third, their approach is not \textit{interpretable} regarding the new solver performance since we are already missing the feature to \textit{rank} the new solver among existing ones.

\subsubsection{Incremental benchmark-set-selection.}
Matricon et. al.~\cite{MatriconAFSH21} present an incremental benchmark-set-selection approach.
Given a pair of SAT solvers, they want to iteratively select a subset of instances to decide with high confidence which of the two solvers is better.
The selection of instances is dependent on the choice of the solvers to distinguish.
They calculate a scoring metric for all instances (that have not been selected yet), run the experiment with the highest score, and update the confidence. If the confidence exceeds a threshold (e.g., 95\%), they stop iterating.
Otherwise, the aforementioned steps are repeated until the desired confidence level is reached or the pool of instances is exhausted.
Most of the proposed strategies in~\cite{MatriconAFSH21} are instance-based methods, i.e., no model of the solver runtimes is built.

For the scoring metric, they discuss (1) a random strategy as a baseline, (2) a discrimination strategy based on the work of Gent et. al.~\cite{GentHJKMNN14}, (3) a variance-based strategy (select instance with the highest variance in its distribution of running times), (4) an information-gain-based method, and (5) a feature-distance-based strategy.
To update the confidence level after each experiment, they discuss using a fixed subset size, a Wilcoxon test, and a distribution-based method. Their experiments show that random instance-selection with a Wilcoxon test as the stopping criterion performs consistently well across different data sets in the sense that it belongs to the Pareto-front or is close to it in all cases.

Regarding our desired features of adaptive benchmark-set-selection (Tab.~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e., it \textit{ranks} solvers, optimizes for \textit{runtime}, is \textit{feature-based}, and uses \textit{incremental non-random} sampling.
Although they provide a confidence metric for their prediction, which makes this approach partially \textit{interpretable}, it is not clear how similar two given solvers are or on which instances they behave similarly, as the prediction is solely binary.
Moreover, the major shortcoming lies within it being not \textit{scalable} with the number solvers.
As they compare only pairs of solvers, scaling with more and more solvers and ideas becomes an issue as there is a quadratic amount of pairs to distinguish. Since a new solver idea is seldomly best at first try but requires more or less tweaking, it is desired to rank a new solver in a way that scales linearly with the number of existing approaches.
In this way, a new solver idea can be discarded if it performs poorly across the board or may be further tweaked if it shows promising results at least in some cases.

\subsubsection{Active learning (AL).}
The posed \textit{new solver problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \textit{new user problem}~\cite{RubensESK15}.
The classic exploration-exploitation dilemma arises since, on the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.

\begin{figure}[tbp!]
  \centering
  \begin{tabular}[c]{ccc}
  \begin{subfigure}[b]{0.25\textwidth}
  \centering
  \resizebox{!}{2.75cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {Training Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {$x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {Approx. Function};
  \node at (-2.8,0.25) {$f$};
  \end{tikzpicture}
  }
  \caption{Passive}
  \label{fig:passive}
  \end{subfigure}
  &
  \begin{subfigure}[b]{0.35\textwidth}
  \centering
  \resizebox{!}{2.75cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {Training Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {$x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {Approx. Function};
  \node at (-2.8,0.25) {$f$};
  \node[right] at (-1.2,2.4) {Pool};
  \draw[->, densely dashed] (-2.4,1.2) arc (-90:0:0.8);
  \draw[->, densely dashed] (-1.6,2.8) arc (0:90:0.8);
  \node[right] at (-1.6,3.135) {$x, ?$};
  \draw  (-1.6,2.4) ellipse (0.4 and 0.3);
  \draw[fill=black]  (-1.725,2.5) ellipse (0.07 and 0.07);
  \draw[fill=black]  (-1.625,2.25) ellipse (0.07 and 0.07);
  \draw[fill=black]  (-1.425,2.4) ellipse (0.07 and 0.07);
  \end{tikzpicture}
  }
  \caption{Active Pool-based}
  \label{fig:activepool}
  \end{subfigure}
  &
  \begin{subfigure}[b]{0.35\textwidth}
  \centering
  \resizebox{!}{2.75cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {Training Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {$x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {Approx. Function};
  \node at (-2.8,0.25) {$f$};
  \node[right] at (-1.1,2.4) {GAN};
  \node[right] at (-1.3,3.135) {$x, ?$};
  \draw[->, densely dashed] (-2.4,1.2) arc (-90:90:1.2);
  \end{tikzpicture}
  }
  \caption{Active Synthesis-based}
  \label{fig:activesynth}
  \end{subfigure}
  \end{tabular}

>>>>

\begin{abstract}
  Experimentation is one of the main phases in algorithm engineering, including the propositional-sat\-is\-fia\-bi\-li\-ty problem (SAT).
  While the field of SAT-solver benchmarking is well established, traditional bench\-mark-set-selection approaches do not optimize for benchmark runtime.
  Benchmark-set selection is the process of choosing a subset of instances out of a pool to be representative and reliable in discriminating SAT solvers.
  This paper presents an active-learning-based benchmark-set-selec\-tion approach that solves the new-solver problem:
  Given a new solver, we want to find its ranking amidst its competitors with as little runtime as possible while still giving accurate ranking predictions.
  We can determine a new solver's rank regarding the SAT Competition 2022 Anniversary Track instances in about \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset.
  We achieve this while still maintaining a prediction accuracy of about \SI{92}{\%} regarding the PAR-2 ranking of this solver if it were evaluated on the complete dataset.
  Additionally, we discuss the importance of instance families in the selection process. 
  Our tool provides the engineer of propositional-satisfiability solvers with a faster way of determining a new SAT-solver's performance in comparison to sampling the benchmark set by hand.

@@

\section{Introduction}
\label{sec:intro}
One of the main phases of algorithm engineering is experimentation.
This is also true for the pro\-po\-si\-tio\-nal-sat\-is\-fia\-bi\-li\-ty problem (SAT), the archetypal $\mathcal{NP}$-complete problem.
While state-of-the-art solvers embark upon more and more problem instances, SAT-solver evaluation is steadily getting more costly.
Usually, solvers submitted to competitions like the annual SAT~Competition~\cite{FroleyksHIJS21,sat2022} themselves precede heavy experimentation and are the remnant of a multitude of ideas and considerations.
While the field of SAT-solver benchmarking is well established, traditional benchmark-set-selection approaches~\cite{Gelder11,HoosKSS13} do not optimize for benchmark runtime.
Their goal is to select a representative and robust benchmark set for discriminating solvers reliably.
Within our work, we show that we can arrive at the same conclusions with high accuracy in only a fraction of the time that would be needed to evaluate the full benchmark set.
To be more specific, we address the following problem:

@@

We want to find a scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ assigning each solver a score.
SAT~Competitions~\cite{FroleyksHIJS21} typically rely on the PAR-2 score, i.e, the average runtime with a penalty term of $2\tau$ for timeouts.
We predict the solver ranking induced by this scoring, without predicting exact solver runtimes.
This allows for both comparing the solvers as well as identifying solvers with similar behavior.
%JB: is there any further behavior we can deduce about a solver scoring than a performance ranking? similar behavior could be something like two solvers being fast/slow on the same instances, but that cannot be deduced from overall score

This current article is proposing a novel active-learning-based benchmark-set-selection approach.
Our algorithm is unique in that it is the only one, to the best of our knowledge, that has all of the following features (also refer to Tab.~\ref{tab:requirements}).
Our benchmarking process outputs a \textit{ranking} to compare the new solver against the existing set of solvers as described above.
We optimize the \textit{runtime} that our strategy needs to arrive at its conclusion.
We use problem-instance and runtime \textit{features} $\varphi\!: \mathcal{A} \times \mathcal{I} \rightarrow \mathbb{R}^n$ since such features already have proven to be helpful in classic benchmark-set-selection approaches~\cite{HoosKSS13} as well as in portfolio selection~\cite{CollauttiMMO13,NgokoCT19}.
%JB: I think we can remove this criterion, as it also is satisified by all competitors (and is commented out in table; if we keep criterion, we should also show it in table)
Moreover, we select instances \textit{non-randomly} and \textit{incrementally}.
By doing so, we have a greater impact on the properties of the benchmarking approach such as its required runtime.
Sampling instances incrementally means taking runtime information from already done experiments into account when choosing the next.
Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \textit{scoring function} that shows by which margin a solver is worse and how similar it is to existing approaches (\textit{interpretability}).
%JB: we should rename "interpretability" criterion, term can have lots of different meanings and our specific meaning is not very intuitive
Finally, our strategy's cost \textit{scales linearly} with the cardinality of the set of existing solvers $\mathcal{A}$ to compare against and the instances $\mathcal{I}$ within our pool of possible benchmark instances.
We avoid comparing pairs of solvers.
This would scale quadratically.
%JB: need to discuss this; log-linear effort might be sufficient if binary-search-/recursive-sorting-like procedure used

To sum up our contribution, we take an existing benchmark set, i.e., the SAT Competition~2022 Anniversary Track~\cite{sat2022}, and predict a new solver's PAR2 score on this dataset with only a fraction of the runtime that would be needed to evaluate this solver on the complete set of instances.
%JB: I think we should expand a bit more on experimental design (how many instances/solvers, what do we compare), and the key results; I think it's okay if the intro becomes a bit longer (and in turn, the overview of the approach's features can be cut a bit)

For reproducibility, all our source code\footnote{temporary, anonymized version for review: \url{xxx}} and and data are available on GitHub\footnote{temporary, anonymized version for review: \url{xxx}}.

@@

Tab.~\ref{tab:requirements} shows a concise overview of existing approaches with a related goal.
%JB: we can probably cut this part a bit, as we discuss weakness of individual approaches later anyway
While traditional benchmark-set-selection~\cite{HoosKSS13} lacks the aspect of runtime optimization, related work within algorithm configuration~\cite{HutterHL11} is concerned with finding better configurations fast.
However, algorithm configuration is not suited for ranking.
There is already existing work employing active learning for SAT-solver discrimination~\cite{MatriconAFSH21}.
This approach has some major shortcomings though.
Finally, our approach seeks to provide all stated features.
Note that all approaches already make use of problem-instance features as it showed to be beneficial in past work~\cite{HoosKSS13,CollauttiMMO13,NgokoCT19}.
%JB: as commented above, we should either remove this criterion or show it in the table

\subsubsection{Traditional benchmark-set-selection.}
%JB: I think we should do some further research on static benchmark-set selection approaches (maybe in SMT, CP, etc.)
Recent studies~\cite{abs-2107-07002,NiesslHWCB22} show that selecting a benchmark set in general is a difficult endeavor for multiple reasons.
Dehghani~et.~al.~\cite{abs-2107-07002} describe how a biased benchmarking set can easily lead to fallacious interpretations.
Moreover, Nießl~et.~al.~\cite{NiesslHWCB22} state that benchmark-set design has many movable parts, e.g., performance measures, aggregation, and handling of missing values.
Questionable research practices might modify those design elements a-posteriori to fit expectations, leading to biased results.

Hoos et. al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting a set of SAT benchmark instances, for instance for a competition.
The selection strategy presented is static, i.e., it does not depend on the solvers to distinguish.
Desirable properties are a variety of instances (avoids over-fitting of solver approaches), adapted instance hardness (not too easy but also not too hard; solvers cannot be distinguished if too many time-out or finish immediately), and no duplicates (having duplicates gives those instances more weight).
To avoid having too similar instances, they make use of a distance-based approach using the SATzilla features~\cite{XuHHL08,features}.
Those features comprise general instance-size features and graph-representation features among others.
%JB: we should probably cut the discussion of each approach's criteria to save space, maybe only mention the unsatisifed ones
Their approach \textit{ranks} solvers by providing its PAR-2 score, is \textit{feature-based}, and is \textit{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
Also, their approach is \textit{interpretable} in the sense that the obtained PAR-2 score induces a distance between solvers.
However, they do not optimize for benchmark \textit{runtime} and select instances \textit{randomly}, apart from constraints on the instance hardness and feature distance.

@@

SMAC~\cite{HutterHL11} tries to find a close-to-optimal configuration for a given SAT-solver on a given set of instances.
%JB: discussion of SMAC is too long, we should cut it a bit
Although this task is different from our goal, e.g., we do not need to navigate configuration space, they use a similar framework.
There are mainly two similarities.
First, both approaches are model-based.
While SMAC tries to gradually build a model of the solver's performance as a function of the configuration parameters, we want to gradually build a model of a new solver's performance based on already known runtimes of instances and runtime information on already known solvers.
Second, both approaches employ some kind of selection strategy that is based upon the previously mentioned model.
Theirs selects configurations rather than instances though.
Furthermore, Hutter et. al.~\cite{HutterHL11} use an approach that makes use of exploitation (what configuration parameters are already good) and exploration (parameter space with little data points) to select the next configuration for which to acquire runtime information. SMAC's configuration sampling has also proven to be better than random sampling in all tested scenarios and has even proven to be significantly better in most cases~\cite{HutterHL11}.

While this approach considers the configuration's \textit{runtime}, is \textit{feature-based}, and \textit{scales linearly} by having a fixed time budget, it is not capable of the other three features listed in Tab.~\ref{tab:requirements}.
First, an algorithm configurator cannot be used to \textit{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
If the new solver does not belong to this category (new ideas are seldomly the best approach right from the start, i.e., they need some tweaking), we do not have any information on how this solver performed.
%JB: "this category" = category of most promising solvers? not 100% clear
Second, while using a model-based selection strategy to sample configurations, instance sampling is done \textit{randomly}, i.e., without building a model over instances.
And third, their approach is not \textit{interpretable} regarding the new solver performance since we are already missing the feature to \textit{rank} the new solver among existing ones.
%JB: well, the internal meta-model of the configurator allows to estimate the solver performance for particular configurations (instead of comparing solvers), which allows a specific performance assessment, so the approach is interpretable in that direction (though for comparison to other solvers, only top performance should be interesting)

\subsubsection{Incremental benchmark-set-selection.}
Matricon et. al.~\cite{MatriconAFSH21} present an incremental benchmark-set-selection approach.
%JB: again, I think there are too many details for describing this single approach
Given a pair of SAT solvers, they want to iteratively select a subset of instances to decide with high confidence which of the two solvers is better.
The selection of instances is dependent on the choice of the solvers to distinguish.
They calculate a scoring metric for all instances (that have not been selected yet), run the experiment with the highest score, and update the confidence. If the confidence exceeds a threshold (e.g., 95\%), they stop iterating.
Otherwise, the aforementioned steps are repeated until the desired confidence level is reached or the pool of instances is exhausted.
Most of the proposed strategies in~\cite{MatriconAFSH21} are instance-based methods, i.e., no model of the solver runtimes is built.

For the scoring metric, they discuss (1) a random strategy as a baseline, (2) a discrimination strategy based on the work of Gent et. al.~\cite{GentHJKMNN14}, (3) a variance-based strategy (select instance with the highest variance in its distribution of running times), (4) an information-gain-based method, and (5) a feature-distance-based strategy.
To update the confidence level after each experiment, they discuss using a fixed subset size, a Wilcoxon test, and a distribution-based method.
Their experiments show that random instance-selection with a Wilcoxon test as the stopping criterion performs consistently well across different data sets in the sense that it belongs to the Pareto front or is close to it in all cases.
%JB: in our evaluation, we should state whether our results are consistent to this or not

Regarding our desired features of adaptive benchmark-set-selection (Tab.~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e., it \textit{ranks} solvers, optimizes for \textit{runtime}, is \textit{feature-based}, and uses \textit{incremental non-random} sampling.
Although they provide a confidence metric for their prediction, which makes this approach partially \textit{interpretable}, it is not clear how similar two given solvers are or on which instances they behave similarly, as the prediction is solely binary.
Moreover, the major shortcoming lies within it being not \textit{scalable} with the number solvers.
As they compare only pairs of solvers, scaling with more and more solvers and ideas becomes an issue as there is a quadratic amount of pairs to distinguish.
%JB: maybe rather log-linear than quadratic, see a prior comment
Since a new solver idea is seldomly best at first try but requires more or less tweaking, it is desired to rank a new solver in a way that scales linearly with the number of existing approaches.
%JB: this general motivation should go there where linearity is introduced as criterion, not here
In this way, a new solver idea can be discarded if it performs poorly across the board or may be further tweaked if it shows promising results at least in some cases.

\subsubsection{Active learning (AL).}
%JB: section does not make clear which of the following approaches we will use (reads more like fundamentals than related work)
The posed \textit{new solver problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \textit{new user problem}~\cite{RubensESK15}.
The classic exploration-exploitation dilemma arises since, on the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
%JB: no sure it is arises here as well; dilemma is typical for searching one optimal solution in continuous space; however, if we find an interesting/useful instance here, I'm not sure if we would continue picking similar instances ("exploting" instead of "exploring") just because we expect them to be useful as well

\begin{figure}[tbp!]
  \centering
  \begin{tabular}[c]{ccc}
  \begin{subfigure}[b]{0.25\textwidth}
  \centering
  \resizebox{!}{2.75cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {Training Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {$x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {Approx. Function};
  \node at (-2.8,0.25) {$f$};
  \end{tikzpicture}
  }
  \caption{Passive}
  \label{fig:passive}
  \end{subfigure}
  &
  \begin{subfigure}[b]{0.35\textwidth}
  \centering
  \resizebox{!}{2.75cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {Training Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {$x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {Approx. Function};
  \node at (-2.8,0.25) {$f$};
  \node[right] at (-1.2,2.4) {Pool};
  \draw[->, densely dashed] (-2.4,1.2) arc (-90:0:0.8);
  \draw[->, densely dashed] (-1.6,2.8) arc (0:90:0.8);
  \node[right] at (-1.6,3.135) {$x, ?$};
  \draw  (-1.6,2.4) ellipse (0.4 and 0.3);
  \draw[fill=black]  (-1.725,2.5) ellipse (0.07 and 0.07);
  \draw[fill=black]  (-1.625,2.25) ellipse (0.07 and 0.07);
  \draw[fill=black]  (-1.425,2.4) ellipse (0.07 and 0.07);
  \end{tikzpicture}
  }
  \caption{Active Pool-based}
  \label{fig:activepool}
  \end{subfigure}
  &
  \begin{subfigure}[b]{0.35\textwidth}
  \centering
  \resizebox{!}{2.75cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {Training Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {$x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {Approx. Function};
  \node at (-2.8,0.25) {$f$};
  \node[right] at (-1.1,2.4) {GAN};
  %JB: does the model need to be a GAN or can it also be a different type of model?
  \node[right] at (-1.3,3.135) {$x, ?$};
  \draw[->, densely dashed] (-2.4,1.2) arc (-90:90:1.2);
  \end{tikzpicture}
  }
  \caption{Active Synthesis-based}
  \label{fig:activesynth}
  \end{subfigure}
  \end{tabular}
